---
title: Walking through RMarkdown
author: Erik Amézquita
date: May 18th, 2020
bibliography: biblio.bib
csl: "csl/harvard-cite-them-right.csl"
output: 
  html_document:
    css: "css/styles2.css"
---

```{r include=F}
library(rmarkdown)
knitr::opts_chunk$set(echo = FALSE)
knitr::opts_chunk$set(include = FALSE)
knitr::opts_chunk$set(engine.path = list(
  python = '/usr/bin/python3',
  ruby = '/usr/bin/ruby'
))
```
## Snapshot: Staff vs Faculty

```{r include=T, echo=T, fig.height=4, fig.align='center'}
source('utils/utils.R')
data <- read.csv('data/college_data.csv')
Y <- "Total.FTE.staff..DRVHR2018."
X <- "Instructional..research.and.public.service.FTE..DRVHR2018."  
ID <- "Institution.Name" 
SZ <- "Total..enrollment..DRVEF2018." 
carnegie <- "Carnegie.Classification..HD2018."
landgrant <- "Land.Grant.Institution..HDNo0Yes8."
sector <- "Sector.of.institution..HDPrivate0Public8."
regression <- linregression_traditional(data,X,Y,carnegie)
summary_reg <- summary(regression)
rsquared <- summary_reg$r.squared
coeffs <- summary_reg$coefficients
```
+ We observe an $R^2$ value of `r rsquared`, which is quite high
+ For every research/instructional employee, we have `r round(coeffs[2,1])` staff employees!

## Use `ggplot2` instead: nicer plots

```{r include=T, echo=T, fig.height=4, fig.align='center'}
foo <- ggplotRegression(data,Y,X,carnegie,sector,'US Doctoral Universities')
foo[[1]]
summary_reg <- summary(foo[[2]])
rsquared <- summary_reg$r.squared
coeffs <- summary_reg$coefficients
```

+ We observe an $R^2$ value of `r rsquared`, which is quite high
+ For every research/instructional employee, we have `r round(coeffs[2,1])` staff employees!

## Use `plotly` to get an interactive plot!

- RMarkdown supports a number of different HTML widgets that breathe life into your data!
- HTML widgets are based on Javascript
    - **Do not** worry about it: **R** got your back.
    - `plotly` works well with `ggplot2`.

```{r include=T, echo=T, out.width='100%', fig.align='center'}
ggplotly(foo[[1]], tooltip = c('id', 'enroll', 'x', 'y'))
```

- Find more information on how to make interactive plots [here](https://www.htmlwidgets.org/showcase_plotly.html).
- _Might_ have issues when converting HTML widgets to PDF.
- Snapshot of widget based on `phantomJS`.

```{r eval=F, echo=T, include=T}
library(webshot)
webshot::install_phantomjs()
```

## Look for other possible correlations

```{r include=T, echo=T, fig.height=4, fig.align='center'}
X <- "Admissions.yield...men..DRVADM2018." 
Y <- "Admissions.yield...women..DRVADM2018." 
foo <- ggplotRegression(data,Y,X,carnegie,sector,'US Doctoral Universities')
foo[[1]]
summary_reg <- summary(foo[[2]])
rsquared <- summary_reg$r.squared
coeffs <- summary_reg$coefficients
```

+ We observe an $R^2$ value of `r rsquared`, which is quite high
+ For every admitted male, we have `r round(coeffs[2,1])` admitted female.