---
title: Walking through RMarkdown
author: Erik Amézquita
date: May 18th, 2020
bibliography: biblio.bib
csl: "csl/harvard-cite-them-right.csl"
output: 
  html_document:
    css: "css/styles2.css"
---

We will examine the file `data/college_data.csv`. The file contains more than 160 different variables for the 395 doctoral universities in the US. The data was obtained from [Integrated Postsecondary Education Data System (IPEDS)](https://nces.ed.gov/ipeds/use-the-data). The variables correspond the most recent available academic year. In most of the cases, this mean 2018-19. However, some variables are from the 2017-18 year.

## Setup

+ We will include a large number of pictures. 
+ We don't want to specify for **every single** code chunk to center and stretch-to-fit every plot. 
+ Thus we **set** those global options from beforehand.

```{r setup}
knitr::opts_chunk$set(fig.align='center')
knitr::opts_chunk$set(out.width='100%')
```

+ Next we **load** the data and our plotting functions from `utils.R`.

```{r data}
source('utils/utils.R')
data <- read.csv('data/college_data.csv')
ID <- "Institution.Name" 
SZ <- "Total..enrollment..DRVEF2018." 
carnegie <- "Carnegie.Classification..HD2018."
landgrant <- "Land.Grant.Institution..HDNo0Yes8."
sector <- "Sector.of.institution..HDPrivate0Public8."
```

+ We will later be able to identify every data point with its institutional name and its total enrollment.
+ The plots can distinguish using two out of three criteria:
    + `carnegie`: Carnegie Classification as of 2018:
        + Very High Research
        + High Research
        + Research
    + `landgrant`: If it is a landgrant institution or not.
    + `sector`: Whether the institution is public or non-for-profit private.

## Snapshot: Staff vs Faculty

```{r staff_vs_faculty}
Y <- "Total.FTE.staff..DRVHR2018."
X <- "Instructional..research.and.public.service.FTE..DRVHR2018."  

regression <- linregression_traditional(data,X,Y,carnegie)

summary_reg <- summary(regression)
rsquared <- summary_reg$r.squared
coeffs <- summary_reg$coefficients
```
+ We observe an $R^2$ value of `r rsquared`, which is quite high
+ For every research/instructional employee, we have `r round(coeffs[2,1])` staff employees!

## Use `ggplot2` instead: nicer plots

```{r }
foo <- ggplotRegression(data,Y,X,carnegie,sector,ID,SZ,'US Doctoral Universities')
foo[[1]]
summary_reg <- summary(foo[[2]])
rsquared <- summary_reg$r.squared
coeffs <- summary_reg$coefficients
```

+ We observe an $R^2$ value of `r rsquared`, which is quite high
+ For every research/instructional employee, we have `r round(coeffs[2,1])` staff employees!

## Use `plotly` to get an interactive plot!

- RMarkdown supports a number of different HTML widgets that breathe life into your data!
- HTML widgets are based on Javascript
    - **Do not** worry about it: **R** got your back.
    - `plotly` works well with `ggplot2`.

```{r }
ggplotly(foo[[1]], tooltip = c('id', 'enroll', 'x', 'y'))
```

- Find more information on how to make interactive plots [here](https://www.htmlwidgets.org/showcase_plotly.html).
- _Might_ have issues when converting HTML widgets to PDF.
- Snapshot of widget based on `phantomJS`.

```{r webshot, eval=F}
library(webshot)
webshot::install_phantomjs()
```

## Look for other possible correlations

```{r enrollment}
X <- "Undergraduate.enrollment..DRVEF2018."  
Y <- "Graduate.enrollment..DRVEF2018."  
foo <- ggplotRegression(data,Y,X,carnegie,sector,ID,SZ,'US Doctoral Universities')
foo[[1]]
summary_reg <- summary(foo[[2]])
rsquared <- summary_reg$r.squared
coeffs <- summary_reg$coefficients
```

+ We observe an $R^2$ value of `r rsquared`, which is low.
+ For every enrolled undergrad, we have `r round(coeffs[2,1])` enrolled grad student.

#### Instead, let's look at full time

```{r full_enrollment}
X <- "Full.time.undergraduate.enrollment..DRVEF2018."  
Y <-  "Full.time.graduate.enrollment..DRVEF2018."  
foo <- ggplotRegression(data,Y,X,carnegie,sector,ID,SZ,'US Doctoral Universities')
foo[[1]]
summary_reg <- summary(foo[[2]])
rsquared <- summary_reg$r.squared
coeffs <- summary_reg$coefficients
```

+ Considering full time enrollment doesn't make the trend clearer.
+ We observe an $R^2$ value of `r rsquared`, which remains low.
+ For every enrolled undergrad, we have `r round(coeffs[2,1])` enrolled grad student.

## What about the number of degrees granted?

+ Are _Very High Research Activity_ institutions producing considerably more academic degrees than the rest of institutions?

```{r phd_production, echo=F}
PhD <- "Number.of.students.receiving.a.Doctor.s.degree..DRVC2018."
MS <- "Number.of.students.receiving.a.Master.s.degree..DRVC2018."
BS <- "Number.of.students.receiving.a.Bachelor.s.degree..DRVC2018."
AS <- "Number.of.students.receiving.an.Associate.s.degree..DRVC2018."
PB <- "Number.of.students.receiving.a.Postbaccalaureate.or.Post.master.s.certificate..DRVC2018."
Cert4 <- "Number.of.students.receiving.a.certificate.of.1.but.less.than.4.years..DRVC2018."
Cert1 <- "Number.of.students.receiving.a.certificate.of.less.than.1.year..DRVC2018."

Men <- "Percent.admitted...men..DRVADM2018."
Women <-  "Percent.admitted...women..DRVADM2018."

GradT <- "Graduate.enrollment..DRVEF2018."
UnderT <- "Undergraduate.enrollment..DRVEF2018."
GradF <- "Full.time.graduate.enrollment..DRVEF2018."
UnderF <- "Full.time.undergraduate.enrollment..DRVEF2018."
Normalized <- "Students.receiving.a.PhD.normalized..DRVC2018."
```
```{r}
p <- ggplotJitter(data, PhD, carnegie, sector, landgrant, ID, SZ)
ggplotly(p, tooltip = c('id', 'enroll', 'y'))
```

+ The VHR institutions have a clear higher output
+ But they also have a larger student body
+ What if we normalize with respect to the number of total enrolled graduate students (we lack the specific number of enrolled PhD students)?
+ Consider then the percentage
$$ \frac{\#(\text{PhD degrees granted})}{\#(\text{Enrolled graduate students})}\times100 $$

```{r}
df <- data.frame(
  data[c(carnegie,PhD,ID,SZ,sector,landgrant)],
  Students.receiving.a.PhD.normalized..DRVC2018. = as.vector(unlist(data[PhD]/data[GradT]))*100
)
Normalized <- "Students.receiving.a.PhD.normalized..DRVC2018."

p <- ggplotJitter(df, Normalized, carnegie, sector, landgrant, ID, SZ)
ggplotly(p, tooltip = c('id', 'enroll', 'y'))
```

## Your Turn

+ Plot other possible linear relationships.
    + Average price vs total enrollment.
    + Instructional vs Research FTEs.
+ Examine bachelor and master output, both raw and normalized.




